\documentclass[nativefonts]{TUD-dissertation2020}

%% Turn off page numbering for the propositions and make the margins on both
%% sides equal and symmetrical.
\geometry{twoside=false}
\pagestyle{empty}
\newcommand{\addmytitle}{\title[\textcolor{tudelft-cyan}{Coupled Local Motion
Planning for Mobile Manipulators}]{Geometric Control}}


\begin{document}

%% Specify the title and author of the thesis. This information will be used on
%% both the English and Dutch side and in the metadata of the final PDF.
\addmytitle
\author{Max}{Spahn}

\begin{center}

{\Large\titlefont\bfseries Propositions}

\bigskip

accompanying the dissertation

\bigskip

%% Print the title.
{\makeatletter
\titlestyle\bfseries\large\@title
\makeatother}

%% Print the optional subtitle.
{\makeatletter
\ifx\@subtitle\undefined\else
    \titlefont\titleshape\@subtitle
\fi
\makeatother}

\bigskip

by

\bigskip

%% Print the full name of the author.
\makeatletter
{\large\titlefont\bfseries\@firstname\ {\titleshape\@lastname}}
\makeatother

\end{center}

\bigskip
\bigskip

\begin{enumerate}

  \item Safety in the sense of collision avoidance is inherently in conflict
    with fast robotic manipulation. (This thesis)
  \item In robotics, robustness is more valuable than optimality and formal
    guarantees. (This thesis)
  \item Optimization fabrics create organic motions. (This thesis)
  \item Warranties on collision avoidance in dynamic environments are not
    realistic, striving for it is thus pointless.
  \item Engineering sciences should embrace failure.
  \item Perception is inaccurate, so the paradigm, sense-plan-act, is prone to
    fail.
  \item Academic publications in robotics are of little use for advancing the
    field without proper software.
  \item In software engineering, comments are a manifestation of failure.
  \item Major advancements in robotics are unlikely to come from academic
    institutions, but rather from companies.
  \item Effective peer-reviewing requires a countable reward system.
  \item Individual career goals harm collaborative research efforts.
  \item Humans quickly get used to a comfort in a way that its removal becomes
    politically impossible.
  \item Effective action to combat climate change require major natural disaster
    for society to understand the urgency.
  \item Advancement in an academic career progressively requires more management
    skills, while these skills are not encouraged in the promotion process.
  \item The older a democracy the more socially segmented its society.
\end{enumerate}

\bigskip
\bigskip

%% Apart from the name and title of the supervisor, the following text is
%% dictated by the promotieregelement.
\begin{center}
These propositions are regarded as opposable and defendable, and have been
  approved as such by the promotor prof.\ dr.\ M.\ Wisse.
\end{center}

\clearpage
{\selectlanguage{dutch}

\begin{center}

{\Large\titlefont\bfseries Stellingen}

\bigskip

behorende bij het proefschrift

\bigskip

%% Print the title.
{\makeatletter
\titlestyle\bfseries\large\@title
\makeatother}

%% Print the optional subtitle.
{\makeatletter
\ifx\@subtitle\undefined\else
    \titlefont\titleshape\@subtitle
\fi
\makeatother}

\bigskip

door

\bigskip

%% Print the full name of the author.
\makeatletter
{\large\titlefont\bfseries\@firstname\ {\titleshape\@lastname}}
\makeatother

\end{center}

\bigskip
\bigskip

\begin{enumerate}

\item Stelling 1.
\item Stelling 2.
\item Stelling 3.
\item Stelling 4.
\item Stelling 5.
\item Stelling 6.
\item Stelling 7.
\item Stelling 8.
\item Stelling 9.
\item Stelling 10.

\end{enumerate}

\bigskip
\bigskip

%% Apart from the name and title of the supervisor, the following text is
%% dictated by the promotieregelement.
\begin{center}
Deze stellingen worden opponeerbaar en verdedigbaar geacht en zijn als zodanig goedgekeurd door de promotor prof.\ dr.\ M.\ Wisse.
\end{center}

}

\end{document}

