\begin{abstract}
  %Short overview of what needs to be done, more details in
  %the individual sections. Don't forgot to put your name in
  %the author list ;).

  %What about the voice-over? In the end, we didn't really
  %use it much but it might still be useful to show it?

  %\MS{Write abstract}

  %\JAM{Overall feedback would be greatly apprectiated.}

  %\CP{Finish decision making, would you be ready to help with a nice video cutting? Combining staff from the hummus video and from the actual demo in Zandam maybe?}


  % Should be resolved now
  % \CS{Do you want to add more specs on the NUC? You also have one comment left
  % in the perception which i think is resolved already.}


  %\CH{We have taken some decision on what to show in the results. Maybe, you can  improve the writing or ask me for more/different quantative results.}

  Although autonomous robots have great potential to boost efficiency and throughput across the whole retail chain, they are mostly being deployed in large warehouses and distribution centers. 
  Deploying robots in stores with customers, such as supermarkets, requires substantially more development efforts since they need to safely operate around customers and reliably cope with various uncertainties and disturbances, such as misplaced products.
  We present our recent efforts in developing a mobile manipulator platform for order picking in realistic supermarket settings. 
  Our robot platform uses state-of-the-art perception and planning algorithms to robustly pick items in the presence of disturbances. In particular, it successfully demonstrates adaptive decision making and rapid motion planning to quickly replan. Our robot also includes a framework to add new products and to teach new picking maneuvers from demonstration. %address various challenges in supermarkets.
  We validated our robot in a recreated supermarket in our lab and in a test supermarket environment of a large Dutch retailer. 
  Our results show how our robot successfully recovers from various disturbances, including misplaced products, errors in picking, and disturbances from human interaction.
  We also summarize our lessons learned to bring autonomous robots into real retail environments with customers.
\end{abstract}
